Clustering in a hyperbolic model of complex networks.

### 2019-03-08 concentration.tex
finished with section 8

Next, recall the adjusted triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus p_0}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p_0}}\ind{p_2 \in \BallPon{p_0}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal},
\]
and recall the definition of $\Kcal_{C,n}(k_n)$
\[
	\Kcal_{C,n}(k_n) = \left\{p \in \Rcal : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi} \right\}.
\]
Slightly abusing notation, we will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\text{box}}(k, C) = \sum_{p \in \Pcal \cap \Kcal_{C,n}(k_n)} \ind{\text{deg}_{\text{box}}(p) = k} \widetilde{T}_{\text{box}}(p).
\end{equation}
The idea is that the main contribution of triangles of degree $k_n$ to $T_{\text{box}}(k_n)$ is given by $\widetilde{T}_{\text{box}}(k_n, C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} it suffices to show that $\widetilde{T}_{\text{box}}(k_n,C)$ is sufficiently concentrated around its mean. This is done in the following proposition.

@@

We postpone the proof of this proposition till Section~\ref{ssec:concentration_tilde_T} and first use it to prove Proposition~\ref{prop:concentration_local_clustering_P_n}.

@@

First we note that $\ind{p_2 \in \BallPo{p_1} \cap \Rcal} \le \ind{p_2 \in \BallPon{p_1}}$ and hence $\widetilde{T}_{\text{box}}(p) \le T_{\text{box}}(p)$. This implies that
\[
	 \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}} \le c^\ast(k_n; \Gbox). 
\]
and therefore
\[
	\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \Exp{c^\ast(k_n; \Gbox)} 
	- \frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}.
\]
For the expectation of $\widetilde{T}_{\text{box}}(k_n,C)$ we use that 
\[
	\CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
= \binom{k_n}{2} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(p)},
\] to get
\begin{align*}
	\Exp{\widetilde{T}_{\text{box}}(k_n,C)} 
	&= \int_{\Kcal_{C,n}(k_n)} \CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
		\rho_{\text{box}}(y,k_n) f(x,y) \dd x \dd y\\
	&= (1+\smallO{1}) \binom{k_n}{2} \int_{\Kcal_{C,n}(k_n)} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(y)}
		\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) \frac{1}{2} \int_{\Kcal_{C,n}(k_n)} \Exp{\widetilde{T}_{\text{box}}(y)}
			\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) n \binom{k_n}{2} \int_0^\infty P(y) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line is due to Corollary~\ref{cor:adjusted_triangle_counting_P_n}. Since $\Exp{N_{\text{box}}(k_n)} = (1+\smallO{1}n p_{k_n}$ it follows that
\[
	\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	= (1+\smallO{1}) \frac{\int_0^\infty P(y) \alpha e^{-\alpha y} \dd y}{p_{k_n}}
	= (1 + \smallO{1}) \gamma(k_n).
\]
On the other hand, Proposition~\ref{prop:convergence_average_clustering_P_n} implies that $\Exp{c^\ast(k_n; \Gbox)} = (1+\smallO{1})\gamma(k_n)$ and thus we conclude that
\[
	2\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \smallO{s(k_n)}.	
\]

For the remaining term we use Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\text{box}}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\text{box}}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	&= \smallO{\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}}
	= \smallO{s(k_n)},
\end{align*}
which finishes the proof.
\end{proof}

@@

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $\Gbox$ and subsequently the joint neighborhoods of two points $p, p^\prime \in \Rcal$. We perform the analysis in this section. The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

\subsubsection*{Neighborhoods}

We start with analyzing joint neighborhoods in $\Gbox$. Let $p, p^\prime \in \Rcal$. Then we denote by $\Ncal_{\text{box}}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $\Gbox$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by $\Ncal_{\text{box}}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish lower bounds on the expected number of points in the disjoint neighborhood $\Exp{\Ncal_{\text{box}}(p\Delta p^\prime)}$ as well as an asymptotic expression for the expected number points in the joint neighborhood. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.

\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
    \draw node at (-0.2,0.7) {$p$};

    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (1,0.5) {};
    \draw node at (0.8,0.2) {\color{blue}$p^\prime$};

	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
    \draw[domain=2.2840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-1)-0.5});
    \draw[domain=-0.2840:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(1-\x)-0.5});

    \draw[dashed,thick,black,line width=1pt] (-6,2.5835) -- (6,2.5835);

    \draw node at (-7,2.5835) {$h(y)$};
    \draw node at (-7,3.3918) {\color{blue}$h_1(p^\prime)$};
    \draw node at (7,2.7189) {\color{blue}$h_2(p^\prime)$};

    \draw node at (-4,3.5) {\color{blue}$x_1 = x^\prime - e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (4.5,3) {\color{blue}$x_1 = x^\prime + e^{\frac{y^\prime + y_1}{2}}$};
    \draw node at (-4.5,1) {$x_1 = x - e^{\frac{y + y_1}{2}}$};
    \draw node at (2,1.5) {$x_1 = x + e^{\frac{y + y_1}{2}}$};

\end{tikzpicture}
\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
\label{fig:representation_disjoint_neighborhoods_P_n}
\end{figure}~
\begin{figure}[p]
\centering
\begin{tikzpicture}

	%The box \Rcal
	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);

    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
    \draw node at (-0.2,0.7) {$p$};

    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (2.5,0.5) {};
    \draw node at (2.3,0.2) {\color{blue}$p^\prime$};

	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
    \draw[domain=3.7840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-2.5)-0.5});
    \draw[domain=1.2160:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(2.5-\x)-0.5});
    \draw[domain=0:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x+9.5)-0.5});

    \draw[dashed,black,line width=1pt] (-6,2.0055) -- (6,2.0055);

%    \draw node at (-7,2.5835) {$h(y)$};
    \draw node at (-7,3.7801) {\color{blue}$h_1(p^\prime)$};
    \draw node at (7,2.0055) {\color{blue}$h_2(p^\prime)$};

%    \draw[dotted,thick,black] (4.5208,2.0174) -- (4.5208,0);
%    \draw[dotted,thick,black] (4.5208,2.0174) -- (6,2.0174);

>>>>

Next, recall the adjusted triangle count function
\[
	\widetilde{T}_{\text{box}}(p_0) = \sum_{(p_1, p_2) \in \Pcal \setminus p_0}^{\ne} 
		\widetilde{T}_{\text{box}}(p_0,p_1,p_2).
\]
where
\[
	\widetilde{T}_{\text{box}}(p_0,p_1,p_2) = \ind{p_1 \in \BallPon{p_0}}\ind{p_2 \in \BallPon{p_0}}\ind{p_2 \in \BallPo{p_1} \cap \Rcal},
\]
and recall the definition of $\Kcal_{C}(k_n)$
\[
	\Kcal_{C}(k_n) = \left\{y \in \R_+ : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi} \vee 1 \le e^{\frac{y}{2}}
		\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi} \right\},
\]
and write $\Rcal(k_n,C) = [-I_n,I_n] \times \Kcal_{C}(k_n)$ for the part of the box $\Rcal$ with heights in $\Kcal_{C}(k_n)$.
Slightly abusing notation, we will define the corresponding triangle degree function
\begin{equation}\label{eq:def_degree_triangle_count_in_K}
	\widetilde{T}_{\text{box}}(k_n, C) = \sum_{p \in \Pcal \cap \Rcal(k_n,C)} \ind{\text{deg}_{\text{box}}(p) = k_n} \widetilde{T}_{\text{box}}(p).
\end{equation}
The idea is that the main contribution of triangles of degree $k_n$ to the triangle count $T_{\text{box}}(k_n)$ is given by $\widetilde{T}_{\text{box}}(k_n, C)$. Therefore, in order to prove Proposition \ref{prop:concentration_local_clustering_P_n} it suffices to show that $\widetilde{T}_{\text{box}}(k_n,C)$ is sufficiently concentrated around its mean. This last part is done in the following proposition.

@@

We first use this result to prove Proposition~\ref{prop:concentration_local_clustering_P_n}. The remainder of this section is devoted to prove Proposition~\ref{prop:concentration_tilde_T_P_n}. The final proof can be found in Section~\ref{ssec:concentration_tilde_T}.

@@

First we note that $\ind{p_2 \in \BallPo{p_1} \cap \Rcal} \le \ind{p_2 \in \BallPon{p_1}}$ and hence $\widetilde{T}_{\text{box}}(p) \le T_{\text{box}}(p)$. This implies that
\[
	 \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}} \le c^\ast(k_n; \Gbox). 
\]
and therefore
\[
	\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \Exp{c^\ast(k_n; \Gbox)} 
	- \frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}.
\]
For the expectation of $\widetilde{T}_{\text{box}}(k_n,C)$ we use that 
\[
	\CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
= \binom{k_n}{2} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(p)},
\] to get
\begin{align*}
	\Exp{\widetilde{T}_{\text{box}}(k_n,C)} 
	&= \int_{\Rcal(k_n,C)} \CExp{\widetilde{T}_{\text{box}}(p)}{\Dbox(p) = k_n}
		\rho_{\text{box}}(y,k_n) f(x,y) \dd x \dd y\\
	&= (1+\smallO{1}) \binom{k_n}{2} \int_{\Rcal(k_n,C)} \Mu{\BallPon{y}}^{-2} \Exp{\widetilde{T}_{\text{box}}(y)}
		\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) \frac{1}{2} \int_{\Rcal(k_n,C)} \Exp{\widetilde{T}_{\text{box}}(y)}
			\rho_{\text{box}}(y,k_n) \alpha e^{-\alpha y} \dd y\\
	&= (1+\smallO{1}) n \binom{k_n}{2} \int_0^\infty P(y) \rho(y,k_n) \alpha e^{-\alpha y} \dd y,
\end{align*}
where the last line is due to Corollary~\ref{cor:adjusted_triangle_counting_P_n}. In particular, since the last integral is $\bigT{k_n^{-(2\alpha + 1)} s(k_n)}$ we conclude that
\begin{equation}\label{eq:scaling_exected_tilde_T}
	\Exp{\widetilde{T}_{\text{box}}(k_n,C)} = \bigT{n k_n^{-(2\alpha - 1)}s(k_n)}. 
\end{equation}

Since $\Exp{N_{\text{box}}(k_n)} = (1+\smallO{1}n p_{k_n}$ it follows that
\[
	\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	= (1+\smallO{1}) \frac{\int_0^\infty P(y) \alpha e^{-\alpha y} \dd y}{p_{k_n}}
	= (1 + \smallO{1}) \gamma(k_n).
\]
On the other hand, Proposition~\ref{prop:convergence_average_clustering_P_n} implies that $\Exp{c^\ast(k_n; \Gbox)} = (1+\smallO{1})\gamma(k_n)$ and thus we conclude that
\[
	2\Exp{\left|c^\ast(k_n; \Gbox) - \frac{\widetilde{T}_{\text{box}}(k_n,C)}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}\right|}
	= \smallO{\gamma(k_n)} = \smallO{s(k_n)}.	
\]

For the remaining term we use H\"{o}lder's inequality and Proposition \ref{prop:concentration_tilde_T_P_n} to obtain
\begin{align*}
	\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}
	&\le \left(\Exp{\widetilde{T}_{\text{box}}(k_n,C)^2} 
		- \Exp{\widetilde{T}_{\text{box}}(k_n,C)}^2\right)^{\frac{1}{2}}\\
	&= \smallO{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}.
\end{align*}
This implies
\begin{align*}
	\frac{\Exp{\left|\widetilde{T}_{\text{box}}(k_n,C) - \Exp{\widetilde{T}_{\text{box}}(k_n,C)}\right|}}
		{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}
	&= \smallO{\frac{\Exp{\widetilde{T}_{\text{box}}(k_n,C)}}{\binom{k_n}{2}\Exp{N_{\text{box}}(k_n)}}}
	= \smallO{s(k_n)},
\end{align*}
which finishes the proof.
\end{proof}

@@

To prove Proposition~\ref{prop:concentration_tilde_T_P_n} we need to understand the joint degree distribution in $\Gbox$. This subsequently requires us to compute the joint neighborhoods in $\Gbox$ of two points $p, p^\prime \in \Rcal$. We perform the analysis in this section. We start with a general result for near independent Poisson random variables.

\begin{lemma}\label{lem:near_independence_poisson}
Let $k_n \to \infty$ and $X_1 = \Po(\lambda_1(n))$, $X_2 = \Po(\lambda_2(n))$ and $Y = \Po(\lambda_3(n))$, be three Poisson random variables where $\lambda_3(n) = \bigO{k_n^{1-\varepsilon}}$, for some $0 < \varepsilon < 1$ and for some $C > 0$,
\[
	k_n - C\sqrt{k_n \log(k_n)} \le \lambda_i(n) + \lambda_3(n) \le k_n + C\sqrt{k_n \log(k_n)},
\] 
for $i = 1,2$.Then, as $n \to \infty$
\[
	\Prob{X_1 + Y = k_n, X_2 + Y = k_n} = (1+\smallO{1}) \Prob{X_1 + Y = k_n}\Prob{X_2 + Y = k_n}.
\]
\end{lemma}

\begin{proof}
First we write
\begin{align*}
	\Prob{X_1 + Y = k_n, X_2 + Y = k_n} 
	&= \sum_{t = 0}^\infty \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t}.
\end{align*}
Now fix a $C_1 > 0$ and define the set
\[
	A_n := \left\{t \in \R_+ \, : \, \lambda_3(n) - C_1\sqrt{k_n^{1-\varepsilon}\log(k_n)} \le t 
	\le \lambda_3(n) + C_1\sqrt{k_n^{1-\varepsilon}\log(k_n)} \right\}.
\]
Then by a Chernoff bound (c.f. \eqref{eq:def_chernoff_bound_poisson})
\begin{align*}
	&\hspace{-20pt}\sum_{t \in \R_+ \setminus A_n} \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t}\\
	&\le \Prob{Y > \lambda_3(n) + C_1\sqrt{k_n^{1-\varepsilon}\log(k_n)}} 
		+ \Prob{Y < \lambda_3(n) - C_1\sqrt{k_n^{1-\varepsilon}\log(k_n)}}\\
	&= \Prob{\left|\Po(\lambda_3(n)) - \lambda_3(n)\right| > C_1\sqrt{k_n^{1-\varepsilon}\log(k_n)}} 
		\le 2 k_n^{-\frac{C_1}{4}}.
\end{align*}
and hence
\[
	\Prob{X_1 + Y = k_n, X_2 + Y = k_n} 
	= \sum_{t \in A_n} \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t} + \bigO{k_n^{-(1+C_1^2)/2}}.
\]
Next, for $i=1,2$ we have by assumption on $\lambda_i(n) + \lambda_3(n)$ that
\begin{align*}
	\Prob{X_i + Y = k_n} &\ge \frac{\left(k_n + C\sqrt{k_n \log(k_n)}\right)^{k_n}}{k_n!} 
		e^{-\left(k_n + C\sqrt{k_n \log(k_n)}\right)}\\
	&\ge e^{-1} k_n^{-1/2} \left(1 + C\sqrt{\frac{\log(k_n)}{k_n}}\right)^{k_n} e^{-C\sqrt{k_n \log(k_n)}}\\
	&\ge e^{-1} k_n^{-1/2} e^{k_n \log\left(1 + C\sqrt{\frac{\log(k_n)}{k_n}}\right) -C\sqrt{k_n \log(k_n)}}\\
	&\ge e^{-1} k_n^{-1/2} e^{ -\frac{C^2}{2} \log(k_n)} = e^{-1} k_n^{-\frac{1+C^2}{2}}.
\end{align*}
where we also used that $\log(1+x) \ge x - x^2/2$, for $0\le x \le 1$ and $k_n! \ge e \sqrt{k_n} k_n^{k_n} e^{-k_n}$. Hence, by taking $C_1 > 4(1 + C^2)$ we get that $k_n^{-\frac{C_1}{4}} = \smallO{\Prob{X_1 + Y = k_n}\Prob{X_1 + Y = k_n}}$. It remains to show that 
\begin{align*}
	&\hspace{-20pt}\sum_{t \in A_n} \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t}\\
	&= (1 + \smallO{1}) \Prob{X_1 + Y = k_n} \Prob{X_2 + Y = k_n},
\end{align*}

For this take any $s \in A_n$ so that $|t-s| \le 2 C_1 \sqrt{k_n^{1-\varepsilon}\log(k_n)}$ and note that there exists a $\delta_n$ satisfying $|\delta_n| \le 2C \sqrt{k_n \log(k_n)}$, for $n$ large enough, such that $k_n - t = \lambda_1(n) + \delta_n$. It then follows that, uniformly in $t,s$ and $\delta_n$, as $n \to \infty$
\begin{align*}
	\frac{\Prob{X_2 = k_n - t}}{\Prob{X_2 = k_n - s}}
	&= \frac{\Prob{X_2 = k_n - t}}{\Prob{X_2 = k_n - t - (s - t)}}\\
	&= \frac{(k_n - t - (s-t))!}{(k_n - t)!} \lambda_1(n)^{s - t} \\
	&\sim (k_n - t - (s-t))^{-(s-t)} \lambda_1(n)^{s-t} \\
	&= \left(\lambda_1(n) + \delta_n - (s-t)\right)^{-(s-t)}\lambda_1(n)^{s-t}\\
	&= \left(1 + \frac{\delta_n -(s-t)}{\lambda_1(n)}\right)^{s-t}\\
	&\sim e^{\frac{(s-t)\delta_n}{\lambda_1(n)}} e^{-\frac{(s-t)^2}{\lambda_1(n)}} \sim 1,
\end{align*}
where the last line follows since both $\frac{(s-t)\delta_n}{\lambda_1(n)} \to 0$ and $\frac{(s-t)^2}{\lambda_1(n)} \to 0$ as $n \to \infty$. In particular,
\[
	\Prob{X_2 = k_n - t} = (1+\smallO{1}) \Prob{X_2 = k_n - s},
\] 
uniformly for all $t, s \in A_n$ and therefore, since
\[
	1 = \sum_{s = 0}^\infty \Prob{Y = s} = (1+\smallO{1}) \sum_{s \in A_n} \Prob{Y = s},
\]
we conclude that
\begin{align*}
	&\hspace{-20pt}\sum_{t \in A_n} \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t}\\
	&= (1+\smallO{1}) \sum_{t \in A_n} \Prob{X_1 = k_n - t}\Prob{X_2 = k_n - t}\Prob{Y = t} \sum_{s \in A_n} \Prob{Y = s}\\
	&= (1 + \smallO{1}) \sum_{t \in A_n} \Prob{X_1 = k_n - t}\Prob{Y = t} 
		\sum_{s \in A_n} \Prob{X_2 = k_n - s} \Prob{Y = s}\\
	&= (1 + \smallO{1}) \Prob{X_1 + Y = k_n} \Prob{X_2 + Y = k_n},
\end{align*}
from which the result follows.
\end{proof}

To see how this lemma can be applied to analyze the joint degree distribution in $\Gbox$, fix two points $p,p^\prime \in \Rcal$ and denote by 
\[
	\rho_{\text{box}}(p,p^\prime,k,k^\prime) 
	:= \Prob{\Po\left(\Mu{\BallPon{p}}\right) = k, \Po\left(\Mu{\BallPon{p^\prime}}\right) = k^\prime}.
\]
the joint degree distribution. Then if we define,
\begin{align*}
	X_1(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p}\setminus \BallPon{p^\prime}}\right),\\
	X_2(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p^\prime}\setminus \BallPon{p}}\right),\\
	Y(p,p^\prime) &:= \Po\left(\Mu{\BallPon{p} \cap \BallPon{p^\prime}}\right)
\end{align*}
it follows that
\[
	\rho_{\text{box}}(p,p^\prime,k_n,k_n) = \Prob{X_1(p,p^\prime) + Y(p,p^\prime) = k_n, X_2(p,p^\prime) + Y(p,p^\prime) = k_n}.
\]
Now, if $y,y^\prime \in \Kcal_{C}(k_n)$ the three Poisson random variables defined above satisfy the condition of Lemma~\ref{lem:near_independence_poisson} regarding the sum $\lambda_i(n) + \lambda_3(n)$. Therefore, if in addition $\Mu{\BallPon{p} \cup \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon}}$, for some $0 < \varepsilon < 1$, we have that 
\[
	\rho_{\text{box}}(p,p^\prime,k_n,k_n) = (1+\smallO{1})\rho_{\text{box}}(p,k_n)\rho_{\text{box}}(p^\prime,k_n). 
\]

To make this more precise, we define, for any $0 < \varepsilon < 1$, the following set
\begin{equation}\label{eq:def_joint_degree_set_E_growing_k}
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Rcal \times \Rcal
		\, : \, y,y^\prime \in \Kcal_{C}(k_n) \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}. 
\end{equation}
We will show (see Corollary~\ref{cor:expected_common_neighbors_Ecal_set}) that for all $(p,p^\prime) \in \mathcal{E}_\varepsilon(k_n)$ it holds that $\Mu{\BallPon{p} \cap \BallPon{p^\prime}} = \bigO{k_n^{1-\varepsilon}}$ and hence the joint degree distribution factorizes on this set. We will use this set later in Section~\ref{ssec:concentration_tilde_T} to prove Proposition~\ref{prop:concentration_tilde_T_P_n}. The main idea behind the above result is that if $p$ and $p^\prime$ are sufficiently separated in the $x$-direction, then the overlap of their neighborhoods $\BallPon{p} \cap \BallPon{p^\prime}$ is of smaller order than $\Mu{\BallPon{p}} + \Mu{\BallPon{p^\prime}}$. We shall therefore proceed with analyzing the joint neighborhoods in $\Gbox$.
%
%
%The following result shows that for any two points in the set
%
%The two main results, which will be the crucial technical ingredients for the proof of Proposition~\ref{prop:concentration_tilde_T_P_n}, are Lemma~\ref{lem:joint_degree_factorization} and Lemma~\ref{lem:joint_degree_distribution_shift}.

\subsubsection*{Joint neighborhoods}

Let $p, p^\prime \in \Rcal$ and denote by
%Then we denote by $\Ncal_{\text{box}}(p\Delta p^\prime)$ the number of disjoint neighbors of $p$ and $p^\prime$ in $\Gbox$, i.e. those points that belong to either $\BallPon{p}$ or $\BallPon{p^\prime}$ but not to their intersection. In addition we denote by 
$\Ncal_{\text{box}}(p,p^\prime)$ the number of joint neighbors of $p$ and $p^\prime$. We shall establish an upper bound on the expected number of joint neighbors when $p$ and $p6\prime$ are sufficiently separated. Observe that $\Exp{\Ncal_{\text{box}}(p,p^\prime)} = \Mu{\BallPon{p} \cap \BallPon{p^\prime}}$.

%as well as an asymptotic expression for the expected number points in the joint neighborhood. For this we will distinguish between the cases where the distance between the $x$-coordinates of $p$ and $p^\prime$ is small or large. Figure~\ref{fig:representation_disjoint_neighborhoods_P_n}-\ref{fig:representation_disjoint_neighborhoods_P_n_3} show the different situations that occur.
%\begin{figure}[p]
%\centering
%\begin{tikzpicture}
%	%The box \Rcal
%	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
%    \draw node at (-0.2,0.7) {$p$};
%    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (1,0.5) {};
%    \draw node at (0.8,0.2) {\color{blue}$p^\prime$};
%	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
%    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
%    \draw[domain=2.2840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-1)-0.5});
%    \draw[domain=-0.2840:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(1-\x)-0.5});
%    \draw[dashed,thick,black,line width=1pt] (-6,2.5835) -- (6,2.5835);
%    \draw node at (-7,2.5835) {$h(y)$};
%    \draw node at (-7,3.3918) {\color{blue}$h_1(p^\prime)$};
%    \draw node at (7,2.7189) {\color{blue}$h_2(p^\prime)$};
%    \draw node at (-4,3.5) {\color{blue}$x_1 = x^\prime - e^{\frac{y^\prime + y_1}{2}}$};
%    \draw node at (4.5,3) {\color{blue}$x_1 = x^\prime + e^{\frac{y^\prime + y_1}{2}}$};
%    \draw node at (-4.5,1) {$x_1 = x - e^{\frac{y + y_1}{2}}$};
%    \draw node at (2,1.5) {$x_1 = x + e^{\frac{y + y_1}{2}}$};
%\end{tikzpicture}
%\caption{Schematic representation of the neighborhoods of $p$ and $p^\prime$ in $\Gbox$ when $|x-x^\prime| \le e^{\frac{y + y^\prime}{2}}$ used for the proof of Lemma~\ref{lem:disjoint_neighbors_P_n} and Lemma~\ref{lem:common_neighbors_Pcal_n}.}
%\label{fig:representation_disjoint_neighborhoods_P_n}
%\end{figure}~
%\begin{figure}[p]
%\centering
%\begin{tikzpicture}
%	%The box \Rcal
%	\draw[line width=1pt] (-6,0) -- (6,0) -- (6,4) -- (-6,4) -- (-6,0);
%    \draw node[fill, circle, inner sep=0pt, minimum size=5pt] at (0,1) {};
%    \draw node at (-0.2,0.7) {$p$};
%    \draw node[fill,blue, circle, inner sep=0pt, minimum size=5pt] at (2.5,0.5) {};
%    \draw node at (2.3,0.2) {\color{blue}$p^\prime$};
%	\draw[domain=1.6487:6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(\x)-1});
%    \draw[domain=-1.6487:-6,smooth,variable=\x,black,line width=1pt] plot (\x, {2*ln(-\x)-1});
%    \draw[domain=3.7840:6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x-2.5)-0.5});
%    \draw[domain=1.2160:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(2.5-\x)-0.5});
%    \draw[domain=0:-6,smooth,variable=\x,blue,line width=1pt] plot (\x, {2*ln(\x+9.5)-0.5});
%    \draw[dashed,black,line width=1pt] (-6,2.0055) -- (6,2.0055);
%%    \draw node at (-7,2.5835) {$h(y)$};
%    \draw node at (-7,3.7801) {\color{blue}$h_1(p^\prime)$};
%    \draw node at (7,2.0055) {\color{blue}$h_2(p^\prime)$};
%%    \draw[dotted,thick,black] (4.5208,2.0174) -- (4.5208,0);
%%    \draw[dotted,thick,black] (4.5208,2.0174) -- (6,2.0174);
%%    \draw node at (4.5208,-0.5) {$w_x(p,p^\prime)$};
%%    \draw node at (7,2.0174) {$w_y(p,p^\prime)$};
%    \draw node at (0,3.5) {\color{blue}$x_1 = x^\prime + e^{\frac{y^\prime + y_1}{2}}$};
%    \draw node at (-2,1.6) {\color{blue}$x_1 = x^\prime - e^{\frac{y^\prime + y_1}{2}}$};
%    \draw node at (-4.5,1) {$x_1 = x - e^{\frac{y + y_1}{2}}$};
%    \draw node at (2,1.5) {$x_1 = x + e^{\frac{y + y_1}{2}}$};
%\end{tikzpicture}
